---
title: "Data Analysis"
author: "Group20"
documentclass: article
output:
  pdf_document: 
    number_sections: yes
    fig_width: 15
    fig_height: 15
    fig_caption: yes
    fig_crop: no
  word_document: default
geometry: left=2cm,right=2cm,top=2cm,bottom=2cm
header-includes:
  - \usepackage{float}
---

------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

```{r libraries, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
library(ggplot2)
library(dplyr) 
library(moderndive) 
library(gapminder) 
library(skimr) 
library(tidyverse)
library(gridExtra)
library(GGally)
library(formatR)
library(qcc)
library(car)
library(MASS)
library(scales)
```
# Introduction {#sec:intro}
comment here...

# Load Data {#sec:load}
```{r loaddata, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
dataset20 <- read.csv("dataset20.csv")
library(knitr)
dataset20$animal_type<-as.factor(dataset20$animal_type)
dataset20$intake_type<-as.factor(dataset20$intake_type)
dataset20$outcome_type<-as.factor(dataset20$outcome_type)
dataset20$chip_status<-as.factor(dataset20$chip_status)
dataset20$month<-as.factor(dataset20$month)
head20<-head(dataset20)#View the structure of data
library(kableExtra)
head20 %>%
  kable(caption = '\\label{tab:summary} Summary statistics on Dallas animal shelter.', align = 'c') %>%
  kable_styling(latex_option = "hold_position")
```
# Exploratory Data Analysis {#sec:exploratory}

```{r exploratory1, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:histgram} Days stay at shelter"}
dataset20 %>%
  ggplot(aes(x = time_at_shelter)) +
  geom_histogram(binwidth = 5,fill="deep pink",color="white") +
  labs(x = "Time at shelter", y = "Number of days")
```

comment here...

```{r exploratory2, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:animal} Pie charts of different time at shelter of different categorical variables"}
#Plot pie charts of 4 categorical variables

pie1<-ggplot(dataset20, aes(x = "animal_counts", fill =animal_type)) + 
  geom_bar(width = 1) + coord_polar(theta = "y") + 
  geom_text(aes(label = ..count..), stat = "count", position = position_stack(vjust =0.5)) +
  theme_void() + ggtitle("Pie chart of animal type")
pie2<-ggplot(dataset20, aes(x = "outcome_type", fill =outcome_type)) + 
  geom_bar(width = 1) + coord_polar(theta = "y") + 
  geom_text(aes(label = ..count..), stat = "count", position = position_stack(vjust =0.5)) +
  theme_void() + ggtitle("Pie chart of outcome type")
pie3<-ggplot(dataset20, aes(x = "chip_status", fill =chip_status)) + 
  geom_bar(width = 1) + coord_polar(theta = "y") + 
  geom_text(aes(label = ..count..), stat = "count", position = position_stack(vjust =0.5)) +
  theme_void() + ggtitle("Pie chart of chip status")
pie4<-ggplot(dataset20, aes(x = "intake_type", fill =intake_type)) + 
  geom_bar(width = 1) + coord_polar(theta = "y") + 
  geom_text(aes(label = ..count..), stat = "count", position = position_stack(vjust =0.5)) +
  theme_void() + ggtitle("Pie chart of intake type")
library(gridExtra)
grid.arrange(pie1,pie2,pie3,pie4,ncol=2,nrow=2)
```

comment here...

```{r exploratory3, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:animal} Boxplots of different time at shelter of different categorical variables"}
#Create 4 box plots of categorical variables and "time_at_shelter" 
#with color mapped to categorical variables
plot1<-ggplot(data =dataset20, aes(x=animal_type, y=time_at_shelter, fill=animal_type)) +
  geom_boxplot() +
  labs(x = "Animal type", y = "Time at shelter")+ 
  theme(legend.position = "none")
plot2<-ggplot(data =dataset20, aes(x=outcome_type, y=time_at_shelter, fill=outcome_type)) +
  geom_boxplot() +
  labs(x = "Outcome type", y = "Time at shelter")+ 
  theme(legend.position = "none")
plot3<-ggplot(data =dataset20, aes(x=chip_status, y=time_at_shelter, fill=chip_status)) +
  geom_boxplot() +
  labs(x = "Chip status", y = "Time at shelter")+ 
  theme(legend.position = "none")
plot4<-ggplot(data =dataset20, aes(x=intake_type, y=time_at_shelter, fill=intake_type)) +
  geom_boxplot() +
  labs(x = "Intake type", y = "Time at shelter")+ 
  theme(legend.position = "none")
library(gridExtra)
grid.arrange(plot1,plot2,plot3,plot4,ncol=2,nrow=2)
```

comment here...

```{r exploratory5, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:month} Boxplots of time at shelter by month"}
ggplot(data = dataset20, mapping = aes(x = factor(month), y = time_at_shelter)) +
  geom_boxplot(fill = "purple") +
  labs(x = "Month", y = "time_at_shelter",
       title = "Boxplots of time at shelter by month")  +
  scale_x_discrete(labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                              "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
```

From the figure \ref{fig:month}, comment here...

# Application Condition Judgement {#sec:check}
## Check of outliers {#sec:outliers}
```{r outliercheck, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:boxplot} Boxplot of number of days at shelter to check for outliers"}
# detection of outliers
boxplot(dataset20$time_at_shelter,
        col=c('orange'),
        ylab="Days",xlab="Time at shelter")
```

The figure \ref{fig:boxplot} indicates multiple outliers, by looking at the maximum values which is 87 which means that the days the animal spent at the shelter was 87. In this case, the 'outliers' we find are consistent with the actual conditions, thus these values can be retained.

## Check of Overdispersion {#sec:overdispersion}
```{r overdispersion_check, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:overdispersion} Scatterplot of overdispersion detection"}
fit1<-glm(time_at_shelter~animal_type+intake_type+outcome_type+chip_status+month,
          data=dataset20,family=poisson(link = "log")) #build poisson glm model
summary(fit1) #view the coefficients
c<-deviance(fit1)/df.residual(fit1) #calculate the deviance over the degree of residual
c 
ggplot(fit1,aes(x=log(fitted(fit1)),y=log((dataset20$time_at_shelter-fitted(fit1))^2)))+
  geom_point(col="#f46d43") + 
  geom_abline(slope=1, intercept=0, col="#a6d96a", size=1) + 
  ylab(expression((y-hat(mu))^2)) + xlab(expression(hat(mu)))
```

comment here(about the summary)...

The proportion is much greater than 1 From the figure \ref{fig:overdispersion}, this appears that most of the points lie above the line of quality for mean and variance. These both imply that may have overdispersion of the data. Then use the qcc library to test whether the data is overdispersion.

```{r overdispersion_check2, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
qcc.overdispersion.test(dataset20$time_at_shelter,type="poisson")
```

The p-value of significance testing is less than 0.05, further indicating overdispersion in data.


## Check of Multicollinearity {#sec:Multicollinearity}
```{r multicollinearty_check, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
fit2<- lm(time_at_shelter~animal_type+intake_type+outcome_type+chip_status+month,
          data=dataset20) #build the linear model
vif(fit2) %>% #calculate VIF
  kable(caption = '\\label{tab:vif} Variance Inflation Factor table', align = 'c') %>%
  kable_styling(latex_option = "hold_position")
```

From the above table, it lists the VIF of the independent variables. It indicates that the VIF of the variables are all less than 10 which means that the independent variables doesn't exist serious multicollinearity.

# Model fitting {#sec:fitting}
```{r modelfitting, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
#Start by fitting a quasi-Poisson model. 
fit3 <- glm(time_at_shelter~animal_type+intake_type+outcome_type+chip_status+month,family= quasipoisson(link="log"),data=dataset20)
summary(fit3)
```

```{r residual_vs_predicted, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:compare} Comparsion of Standardised Pearson residual plots between regular likelihood and Quasi-likelihood"}
pred<-predict(fit1,type="response")
stand.resid<-rstandard(model=fit1,type="pearson")
par(mfrow=c(1,2),pin=c(6,6))
plot(x=pred,y=stand.resid,xlab="Predicted count",ylab="Standardised Pearson residuals",
     main="Regular likelihood",ylim=c(-2,4))
abline(h=c(-1.5,0,4),lty="dotted",col="red")
pred<-predict(fit3,type="response")
stand.resid<-rstandard(model=fit3,type="pearson")
plot(x=pred,y=stand.resid,xlab="Predicted count",ylab="Standardised Pearson residuals",
     main="Quasi-likelihood",ylim=c(-2,4))
abline(h=c(-1.5,0,4),lty="dotted",col="red")
```
We can see in the plots that most residuals are contained within (-1.5,4) in the quasi-Poisson model, while quite more are outside this range for the original Poisson model.

Next we fit a negative binomial model:
```{r negativebinomial, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
fit4<- glm.nb(time_at_shelter~animal_type+intake_type+outcome_type+chip_status+month,
               data = dataset20)
summary(fit4)
```

To decide between the quasi-Poisson and the negative binomial model, we plot $(y_{\mbox{i}}-\widehat{\mu}_{\mbox{i}})^2$ v $\widehat{\mu}_{\mbox{i}}$ and compare the linear and quadratic fit to see which one of them fits better. 

```{r plot, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:choose} Plot of square residuals v predicted values, quasi-Poisson model (bule line), negative binomial model (red line)"}
res.sq<-residuals(fit1,type = "response")^2
set1 <- data.frame(res.sq,mu.hat=fit1$fitted.values)
fit.lin <- lm(formula = res.sq ~ mu.hat,data = set1)
fit.quad <- lm(formula = res.sq ~ mu.hat + I(mu.hat^2),data = set1)
summary(fit.quad)

plot(set1$mu.hat,y=set1$res.sq,xlab = "Predicted count",ylab="Squared Residul")
curve(expr = predict(fit.lin,newdata = data.frame(mu.hat=x),type="response"),
      col="blue",add = TRUE,lty="solid")
curve(expr = predict(fit.quad,newdata = data.frame(mu.hat=x),type="response"),
      col="red",add = TRUE,lty="dashed")
legend("topleft",legend=c("Linear","Quadratic"),col=c("blue","red"),
      lty=c("solid","dashed"),bty="n")
```
The straight line fit (blue line) indicate that the quasi_Poisson model, and the quadratic fit (red line) would suggest using a negative binomial model.

The quadratic coefficient in the above model has a $p$-value of 0.9115, which is not significant. We don't have any evidence that the negative binomial is better and we go with quasi-Poisson model.

In the original quasi-Poisson model, we detected that the $p$-values for different animal types are all not significant (<0.05), hence, we removed it from the model.

```{r modelselection, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
fit5 <- glm(time_at_shelter ~ intake_type+outcome_type+chip_status+month,family=quasipoisson(link="log"),data=dataset20)
summary(fit5)
```

In the new quasi-Poisson model, we detected that the only one $p$-value for month is significant (<0.05), hence, we removed it from the model.

```{r modelselection2, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
fit6 <- glm(time_at_shelter ~ intake_type+outcome_type+chip_status,family=quasipoisson(link="log"),data=dataset20)
summary(fit6)
fit6 %>%
  coef() %>%
  exp()
```
# Conclusion {#sec:conclusion}
