---
title: "Data Analysis"
author: "Group20"
documentclass: article
output:
  pdf_document: 
    number_sections: yes
    fig_width: 15
    fig_height: 15
    fig_caption: yes
    fig_crop: no
  word_document: default
geometry: left=2cm,right=2cm,top=2cm,bottom=2cm
header-includes:
  - \usepackage{float}
---

------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

```{r libraries, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE}
library(ggplot2)
library(dplyr) 
library(moderndive) 
library(gapminder) 
library(skimr) 
library(tidyverse)
library(gridExtra)
library(GGally)
library(formatR)
library(qcc)
library(car)
```
# Introduction {#sec:intro}
comment here...

# Load Data {#sec:load}
```{r loaddata, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
setwd("D://master//data analysis skills//Group Work2")
dataset20 <- read.csv("dataset20.csv")
library(knitr)
dataset20$animal_type<-as.factor(dataset20$animal_type)
dataset20$intake_type<-as.factor(dataset20$intake_type)
dataset20$outcome_type<-as.factor(dataset20$outcome_type)
dataset20$chip_status<-as.factor(dataset20$chip_status)
dataset20$month<-as.factor(dataset20$month)
head20<-head(dataset20)#View the structure of data
library(kableExtra)
kable(head20,format="latex")%>%
  add_header_above(c(" "=1,"The structure of data"=2," "=1))%>%
  kable_styling(full_width = F)
```
# Exploratory Data Analysis {#sec:exploratory}

```{r exploratory1, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:histgram} Days stay at shelter"}
dataset20 %>%
  ggplot(aes(x = time_at_shelter)) +
  geom_histogram(binwidth = 5,fill="deep pink",color="white") +
  labs(x = "Time at shelter", y = "Number of days")
```
comment here...
```{r exploratory2, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:animal} Different time at shelter of different categorical variables"}
#Plot pie charts of 4 categorical variables
pie1<-ggplot(dataset20, aes(x = "animal_counts", fill =animal_type)) + 
  geom_bar(width = 1) + coord_polar(theta = "y") + 
  geom_text(aes(label = ..count..), stat = "count", position = position_stack(vjust =0.5)) +
  theme_void() + ggtitle("Pie chart of animal type")
pie2<-ggplot(dataset20, aes(x = "outcome_type", fill =outcome_type)) + 
  geom_bar(width = 1) + coord_polar(theta = "y") + 
  geom_text(aes(label = ..count..), stat = "count", position = position_stack(vjust =0.5)) +
  theme_void() + ggtitle("Pie chart of outcome type")
pie3<-ggplot(dataset20, aes(x = "chip_status", fill =chip_status)) + 
  geom_bar(width = 1) + coord_polar(theta = "y") + 
  geom_text(aes(label = ..count..), stat = "count", position = position_stack(vjust =0.5)) +
  theme_void() + ggtitle("Pie chart of chip status")
pie4<-ggplot(dataset20, aes(x = "intake_type", fill =intake_type)) + 
  geom_bar(width = 1) + coord_polar(theta = "y") + 
  geom_text(aes(label = ..count..), stat = "count", position = position_stack(vjust =0.5)) +
  theme_void() + ggtitle("Pie chart of intake type")
library(gridExtra)
grid.arrange(pie1,pie2,pie3,pie4,ncol=2,nrow=2)
```
comment here...
```{r exploratory3, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:animal} Different time at shelter of different categorical variables"}
#Create 4 box plots of categorical variables and "time_at_shelter" 
#with color mapped to categorical variables
plot1<-ggplot(data =dataset20, aes(x=animal_type, y=time_at_shelter, fill=animal_type)) +
  geom_boxplot() +
  labs(x = "Animal type", y = "Time at shelter")+ 
  theme(legend.position = "none")
plot2<-ggplot(data =dataset20, aes(x=outcome_type, y=time_at_shelter, fill=outcome_type)) +
  geom_boxplot() +
  labs(x = "Outcome type", y = "Time at shelter")+ 
  theme(legend.position = "none")
plot3<-ggplot(data =dataset20, aes(x=chip_status, y=time_at_shelter, fill=chip_status)) +
  geom_boxplot() +
  labs(x = "Chip status", y = "Time at shelter")+ 
  theme(legend.position = "none")
plot4<-ggplot(data =dataset20, aes(x=intake_type, y=time_at_shelter, fill=intake_type)) +
  geom_boxplot() +
  labs(x = "Intake type", y = "Time at shelter")+ 
  theme(legend.position = "none")
library(gridExtra)
grid.arrange(plot1,plot2,plot3,plot4,ncol=2,nrow=2)
```
comment here...

```{r exploratory5, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:month} Boxplots of time at shelter by month"}
ggplot(data = dataset20, mapping = aes(x = factor(month), y = time_at_shelter)) +
  geom_boxplot(fill = "purple") +
  labs(x = "Month", y = "time_at_shelter",
       title = "Boxplots of time at shelter by month")  +
  scale_x_discrete(labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                              "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))
```

From the figure \ref{fig:month}, comment here...

# Application Condition Judgement {#sec:check}
## Check of outliers {#sec:outliers}
```{r outliercheck, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:boxplot} Outlier detection"}
# detection of outliers
boxplot(dataset20$time_at_shelter,
        col=c('orange'),
        ylab="Days",xlab="Time at shelter")
```

The figure \ref{fig:boxplot} indicates multiple outliers, by looking at the maximum values which is 87 which means that the days the animal spent at the shelter was 87. In this case, the 'outliers' we find are consistent with the actual conditions, thus these values can be retained.

## Check of Overdispersion {#sec:overdispersion}
```{r overdispersion_check, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
fit1<-glm(time_at_shelter~intake_type+outcome_type+chip_status+month,
          data=dataset20,family=poisson(link = "log")) #build poisson glm model
summary(fit1) #view the coefficients
c<-deviance(fit1)/df.residual(fit1) #calculate the deviance over the degree of residual
c 
```

comment here(about the summary)...

The proportion is much greater than 1, then use the qcc library to test whether the data is overdispersion. 

```{r overdispersion_check2, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
qcc.overdispersion.test(dataset20$time_at_shelter,type="poisson")
```

The p-value of significance testing is less than 0.05, further indicating overdispersion in data.

In this case, replacing family="poisson" by family="quasipoisson" to solve this problem.

## Check of Multicollinearity {#sec:Multicollinearity}
```{r multicollinearty_check, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
fit2<- lm(time_at_shelter~intake_type+outcome_type+chip_status+month,
          data=dataset20) #build the linear model
vif(fit2) #calculate VIF
```

From the above table, it lists the VIF of the independent variables. It indicates that the VIF of the variables are all less than 10 which means that the independent variables doesn't exist serious multicollinearity.

# Model fitting {#sec:fitting}
```{r modelfitting, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
fit3<-glm(time_at_shelter~intake_type+outcome_type+chip_status+month,data=dataset20,
          family=quasipoisson(link = "log")) #build quasipoisson glm model
summary(fit3) #view the coefficients
exp(coef(fit3))

```
comment here...

```{r residual_vs_predicted, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE,fig.align = "center",fig.dim=c(13,8), fig.pos = "H",fig.cap = "\\label{fig:compare} Residual Plots vs. Predicted"}
pred<-predict(fit1,type="response")
stand.resid<-rstandard(model=fit1,type="pearson")
par(mfrow=c(1,2),pin=c(6,6))
plot(x=pred,y=stand.resid,xlab="Predicted count",ylab="Standardised Pearson residuals",
     main="Regular likelihood",ylim=c(-2,4))
abline(h=c(-1.5,0,4),lty="dotted",col="red")
pred<-predict(fit3,type="response")
stand.resid<-rstandard(model=fit3,type="pearson")
plot(x=pred,y=stand.resid,xlab="Predicted count",ylab="Standardised Pearson residuals",
     main="Quasi-likelihood",ylim=c(-2,4))
abline(h=c(-1.5,0,4),lty="dotted",col="red")
```
We can see in the plots that most residuals are contained within (-1.5,4) in the quasi-Poisson model, while quite more are outside this range for the original Poisson model.
